% UG project example file, February 2022
% Do not change the first two lines of code, except you may delete "logo," if causing problems.
% Understand any problems and seek approval before assuming it's ok to remove ugcheck.
\documentclass[bsc,singlespacing,parskip,logo]{infthesis}
\usepackage{ugcheck}

% Include any packages you need below, but don't include any that change the page
% layout or style of the dissertation. By including the ugcheck package above,
% you should catch most accidental changes of page layout though.
\usepackage{microtype} % recommended, but you can remove if it causes problems
\usepackage{quiver}
\usepackage{hyperref}

\begin{document}
\begin{preliminary}

  \title{Towards Strong Relative Pseudomonads}

  \author{Franz Miltz}

  % CHOOSE YOUR DEGREE a):
  % please leave just one of the following un-commented
  %\course{Artificial Intelligence}
  %\course{Artificial Intelligence and Computer Science}
  %\course{Artificial Intelligence and Mathematics}
  %\course{Artificial Intelligence and Software Engineering}
  %\course{Cognitive Science}
  %\course{Computer Science}
  %\course{Computer Science and Management Science}
  \course{Computer Science and Mathematics}
  %\course{Computer Science and Physics}
  %\course{Software Engineering}
  %\course{Master of Informatics} % MInf students

  % CHOOSE YOUR DEGREE b):
  % please leave just one of the following un-commented
  %\project{MInf Project (Part 1) Report}  % 4th year MInf students
  %\project{MInf Project (Part 2) Report}  % 5th year MInf students
  \project{4th Year Project Report}        % all other UG4 students

  \date{\today}

  \abstract{
    Synthetic measure theory uses commutative monads to develop an entirely categorical
    language of measures and integration. This language has proven practically useful
    in the development of higher-order statistical programming languages.

    There is another purely categorical notion of integration: coends. Certain coends
    arise from the presheaf construction, a monad-like structure that fails to be a
    model of synthetic measure theory for several reasons. There are multiple ways
    in which one could attempt to fix these problems.

    In this report, we make a start at defining a strong relative pseudomonad that
    would be suitable as the backbone of an extended synthetic measure theory. At each
    step, we show how the presheaf construction gives rise to the required structure and
    how this structure satisfies the necessary axioms.

    Unfortunately, the strength of a monad does not lend itself to being generalised
    to relative pseudomonads. We are thus neither able to give a complete definition
    of a strong relative pseudomonad nor do we manage to extend synthetic measure theory
    to admit the presheaf construction as a model.
  }

  \maketitle

  \newenvironment{ethics}
  {\begin{frontenv}{Research Ethics Approval}{\LARGE}}
  {\end{frontenv}\newpage}

  \begin{ethics}
    This project was planned in accordance with the Informatics Research
    Ethics policy. It did not involve any aspects that required approval
    from the Informatics Research Ethics committee.

    \standarddeclaration
  \end{ethics}


  \begin{acknowledgements}
    I am incredibly grateful to Dr Ohad Kammar, my supervisor, who started to
    talk to me about category theory well before I began my work on this
    project. Thank you for your patience, your encouragement, and the vast
    amount of knowledge that you have been able to share with me.

    Thanks should also go to Mr Matthias K\"onig. Thank you for setting me on
    the magnificent path towards abstract mathematics. I would not be where I am
    today without you.

    Finally, I would like to express my sincerest appreciation to Mary
    and Fred. You have never failed to put a smile on my face.
  \end{acknowledgements}


  \tableofcontents
\end{preliminary}

\include{thesis/introduction.tex}

\include{thesis/category_theory.tex}

\include{thesis/2categories.tex}

\include{thesis/strong_inclusion_pseudomonads.tex}

\include{thesis/conclusion.tex}

\bibliographystyle{plain}
\bibliography{sources}

%\appendix

%\include{thesis/symmetric_monoidal_categories.tex}

\end{document}
